---
title: "Data Challenge 1"
author: "Leyi Xu"
date: '2022-02-25'
output: 
  html_document:
    code_folding: show
    fig_caption: yes
    highlight: tango
    number_sections: no
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: yes
fig_width: 6
---

<style>
body {text-align: justify}
</style>

# Preliminaries
```{r echo=FALSE, message=FALSE, warning=TRUE}

# ------------------------------------------------------------------------------
options(scipen = 5000)
options(digits.secs = 4)
options(warning = FALSE)


# ------------------------------------------------------------------------------
# packages:

## Default repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos=r)
})

check_pkg <- function(x)
  {
    if (!require(x, character.only = TRUE, quietly = TRUE))
    {
      install.packages(x, dep = TRUE, verbose = FALSE, quiet = TRUE)
        if(!require(x, character.only = TRUE, quietly = TRUE)) stop("Package not found")
    }
}

check_pkg("tidyverse")
check_pkg("sf")
check_pkg("ggplot2")
check_pkg("tmap")
check_pkg("geosphere")
check_pkg("here")


here::i_am("Data Challenge 1_Leyi Xu.Rmd")
```


# 1. How many trips are reported in the data?
```{r load_data, echo = TRUE, warning= FALSE, message = FALSE}
# Load citibike data
citibike <- read.csv("201903-citibike-tripdata.csv")

citibike_num <- nrow(citibike)
```
There are `r citibike_num` trips reported in the data.

# 2. What is the earliest time of the day any trip starts?
```{r earliest_time, echo = TRUE, warning= FALSE, message = FALSE}
library(lubridate)

citibike <- citibike %>%
  # add the column start_time which converts character into a date-time format
  mutate(start_time = as.POSIXct(starttime, format="%Y-%m-%d%H:%M:%OS")) %>%
  # add the column date which extracts year, month and date from date-time
  mutate(start_date = as_date(starttime)) %>%
  # add the column hms which extracts hour, minute and second from date-time
  mutate(start_hms = format(start_time, "%H:%M:%OS"))

# sort the time by order and select the first (earliest) one
earliest_time <- sort(citibike$start_hms)[1]
```
The earliest time of the day any trip starts is `r earliest_time`.

# 3.  How many stations are used?
```{r stations, echo = TRUE, warning= FALSE, message = FALSE}
library(dplyr)

# Collect all the stations used (start stations + end stations)
stations = data.frame(stations = c(citibike$start.station.id, citibike$end.station.id))
stations = stations %>%
  # filter out the null
  filter(!stations == "NULL") %>%
  # delete duplicated stations
  unique()

stations_num <- length(stations$stations)
```
There are `r stations_num` stations used.

# 4. What is the mean number of trips per day?
```{r trips_per_day, echo = TRUE, warning= FALSE, message = FALSE}
# count the number of trips each day
trips <- citibike %>%
  group_by(start_date) %>%
  summarise(count = n())

# calculate the mean number of trips per day
trips_mean_perday <- mean(trips$count)
```
The mean number of trips per day is about `r trips_mean_perday`

# 5. What is the mean number of trips on Sundays?
```{r trips_Sundays, echo = TRUE, warning= FALSE, message = FALSE}
# get the day of week from the date
trips <- trips %>%
  mutate(wd = wday(start_date))

# filter Sundays
trips_sundays <- trips %>%
  filter(wd == 1)

# calculate the mean number of trips on Sundays
trips_mean_sundays <- mean(trips_sundays$count)
```
The mean number of trips on Sundays is about `r trips_mean_sundays`

# 6. On which day did most trips start?
```{r most_trips, echo = TRUE, warning= FALSE, message = FALSE}
# filter the day with most trips
trips_max <- trips %>%
  filter(count == max(count))

# get the date
trips_max_date <- trips_max$start_date
```
Most trips start on `r trips_max_date`.

# 7. That day, how many trips started between 11:00 and 15:00?
```{r trips_11_15, echo = TRUE, warning= FALSE, message = FALSE}
# get the trips started between 11:00 and 15:00 on 2019-03-30
trips_11_15 <- citibike %>%
  filter(start_date == trips_max_date) %>%
  filter(start_hms >= "11:00:00.0000" & start_hms <= "15:00:00.0000")

trips_11_15_num <- nrow(trips_11_15)
```
There are `r trips_11_15_num` trips between 11:00 and 15:00 on `r trips_max_date`.

# 8. What is the duration of the longest trip?
```{r trips_duration, echo = TRUE, warning= FALSE, message = FALSE}
# get the maximum of tripduration and convert seconds to days
duration_longest <- max(citibike$tripduration)/86400
```
The longest trip lasted for about 34 days.

# 9. What is the 99% quantile of trip durations?
```{r quantile_99, echo = TRUE, warning= FALSE, message = FALSE}
quantile_99 <- quantile(citibike$tripduration, probs = .99)/60
```
The 99% quantile of trip durations is about `r quantile_99` minutes.

# 10. Plot a histogram of all trips shorter than 45 minutes.
```{r hist_plot, echo = TRUE, warning= FALSE, message = FALSE}
# get trips shorter than 45 minutes
trips_45 <- citibike %>%
  mutate(tripduration = round(tripduration/60)) %>%
  filter(tripduration < 45)

# plot the histogram
ggplot(trips_45, aes(x = tripduration)) +
  geom_histogram(binwidth = 1) +
  xlab("Trip duration in minutes") + ylab("Number of trips") +
  ggtitle("Trip duration in March")
  
```

# 11. What is the median beeline distance travelled?
```{r median, echo = TRUE, warning= FALSE, message = FALSE}
library(geosphere)

# calculate the travel distance between start and end stations
citibike$dist<-distGeo(data.frame(citibike$start.station.longitude, citibike$start.station.latitude), 
                       data.frame(citibike$end.station.longitude, citibike$end.station.latitude))

# get the median of travel distance
dist_median <- median(citibike$dist)
```
The median distance travelled is `r dist_median`m.

# 12. Generate a sub-sample containing a random 10% of the data, and for that subsample plot the time to ride the (beeline) distance.
```{r plot_time_distance, echo = TRUE, warning= FALSE, message = FALSE}
# get the extreme values in trip duration
extreme_high <- quantile(citibike$tripduration, probs = .75) + 1.5*IQR(citibike$tripduration)

# extreme_low <- quantile(citibike$tripduration, probs = .25) - 1.5*IQR(citibike$tripduration)

# filter out the objects with extreme values in trip duration
citibike_new <- citibike %>%
  filter(tripduration < extreme_high)

# select random 10% of citibike as a sub-sample
citibike_sample <- citibike_new[sample(nrow(citibike_new), nrow(citibike_new)*0.1), ]

# plot the time to distance
ggplot(citibike_sample, aes(x = dist, y = tripduration)) +
  geom_point(size = 0.1) +
  coord_fixed(ratio= 7) +
  xlab("Beeline distance [m]") + ylab("Trip duration [s]") +
  ggtitle("Trip duration by distance covered")
```

# 13. Plot a map indicating the 10 most and least active stations.
```{r plot_stations, echo = TRUE, warning= FALSE, message = FALSE}
# create start stations sf object and end stations sf objects according to their latitudes and longitudes
start_sf = st_as_sf(citibike, coords = c("start.station.longitude", "start.station.latitude"), crs = 4326)
end_sf = st_as_sf(citibike, coords = c("end.station.longitude", "end.station.latitude"), crs = 4326)

# get all stations (start stations and end stations)
stations_all <- st_as_sf(data.frame(id = c(start_sf$start.station.id, end_sf$end.station.id), geometry = c(start_sf$geometry, end_sf$geometry)))

# count how many times each station is used
stations_all_count <- stations_all %>%
  filter(!id == "NULL") %>%
  group_by(id) %>%
  summarise(count = n())

# get the 10 most and least active stations
stations_active <- rbind(stations_all_count[order(stations_all_count$count), ][1:10, ],
                         stations_all_count[order(stations_all_count$count, decreasing = TRUE), ][1:10, ])

# add the column active to indicate each station is active or not
stations_active$active <- "no"
stations_active$active[11:20] <- "yes"

# rescale the count as radius to visualize the 10 most and least active stations
stations_active$radius <- stations_active$count
stations_active$radius[11:20] <- stations_active$count[11:20]/20


# plot the map
tmap_mode("view")

tm_shape(stations_active) + 
  tm_dots(size = "radius", col = "active", palette = "Dark2", alpha = 0.8) +
  tm_basemap("Stamen")
```